### 2017-11-01 Maths Notes/C4/Coordinate Geometry/Coordinate Geometry.tex

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{amssymb}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{mwe}
\usepackage{float}
\usepackage{polynom}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - C4}
\rfoot{Page \thepage}

\newcommand{\R}{\mathbb{R}}
\pgfplotsset{ytick style={draw=none}}
\pgfplotsset{xtick style={draw=none}}

\begin{document}
\begin{center}
\underline{\huge Coordinate Geometry}
\end{center}
\section{Parametric equations}
In parametric equations, coordinates x and y are expressed as functions.\\
$x=f(t)$ and $y=g(t)$ where t is the independent variable (called a parameter)
\subsection{Converting parametric equations into Cartesian form}
\textbf{Example 1}\\
$x=2t$\\
$y=t^2$\\
\\
$t=\dfrac{x}{2}$\\
\\
$y=\frac{1}{4}x^2$\\
\\
\textbf{Example 2}\\
$x=\sin(t)+2$\\
$y=\cos(t)-3$\\
\\
$\sin(t)=x-2$\\
$\cos(t)=y+3$\\
\\
$\sin^2\theta+\cos^2\theta=1$\\
$(x-2)^2+(y+3)^2=1$
\section{Parametric differentiation}
The chain rule can be used to differentiate parametric equations\\
$\dfrac{dy}{dx}=\dfrac{dy}{dt}\times\dfrac{dt}{dx}$\\
This can be used to find the gradient of normals and tangents.\\
\\
\textbf{Example}\\
$x=t^3+t$\\
$y=t^2+1$\\
\\
\textbf{Differentiate the x and y terms}\\
$\dfrac{dy}{dt}=2t$\\
$\dfrac{dx}{dt}=3t^2+1$\\
\\
\textbf{Invert the differential of x to fit the formula}\\
$\dfrac{dt}{dx}=\dfrac{1}{3t^2+1}$\\
\\
\textbf{Multiply together to make one fraction}\\
$\dfrac{dy}{dx}=\dfrac{2t}{3t^2+1}$

\end{document}