### 2017-11-12 Maths Notes/D2/Travelling Salesman Problem/Travelling Salesman Problem.tex
Notes and Timetable update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{enumerate}
\usepackage{amssymb}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{tikz}
\usetikzlibrary{calc}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}	
\newcommand{\cred}[1]{\color{red}#1}
\newcommand{\cblue}[1]{\color{blue}#1}
\newcommand{\cgreen}[1]{\color{green}#1}
\hyphenpenalty=10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - D2}
\rfoot{Page \thepage}
\usepackage{float}

\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

\newcommand{\DrawVLine}[3][]{%
    \begin{tikzpicture}[overlay,remember picture]
        \draw [#1] ($(#2.north)$) -- ($(#3.north)$);
    \end{tikzpicture}%
}%
\newcommand{\DrawLine}[3][]{%
    \begin{tikzpicture}[overlay,remember picture]
        \draw [#1] ($(#2)+(-0.4,0.6ex)$) -- ($(#3)+(0.6,0.6ex)$);
    \end{tikzpicture}%
}%

\begin{document}
\begin{center}
\underline{\huge Travelling Salesman Problem}
\end{center}
\section{Definitions}
The travelling salesman problem looks for a \textbf{walk} that gives the minimum \textbf{tour}.\\
\\
\textbf{Walk} - A finite series of edges so that the end of one vertex is the start of the next\\
\\
\textbf{Tour} - A walk that visits every vertex and returns to the starting vertex
\\
\textbf{Add more information about upper and lower bounds here when you understand it more}
\section{The differences between classical and practical problems}
\textbf{Classical Problem} - Must visit each vertex \textbf{only once} before returning to the start.\\
\textbf{Practical Problem} - Must visit each vertex \textbf{at least once} before returning to the start
\section{Converting a network into a complete network of least distances}
If a network is converted into a complete network of least distances, the classical and practical problem are the same.\\
\\
To create a complete network of least distances, you must ensure the \textbf{triangle inequality} holds for all triangles in the network.\\
\\
\textbf{Triangle inequality:}\\
The longest side of any triangle $\leqslant$ The sum of the two shorter sides\\
\\
In a network where the triangle inequality does not hold, replace the longest arc with the sum of the two shorter ones.
\end{document}