### 2017-11-18 Physics Notes/Practical Notes/Practical Notes.tex
Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{circuitikz}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{relsize}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Physics - Practical Skills}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Practical Skills}
\end{center}
\section{\% error on gradient}
Total \% error = \% error on X + \% error on Y\\
\\
\% error on X = 1 Small Square/Number of squares on triangle\\
\% error on Y = 1 Small Square/Number of squares on triangle\\
\end{document}