### 2017-12-05 Maths Notes/M2/Dynamics/Dynamics.tex
Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{tikz}
\usetikzlibrary{positioning}
\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Dynamics}
\end{center}
\newpage
\begin{center}
\underline{\huge Dynamics Example - F=ma on a slope}
\end{center}
\textit{A car of mass 1500 kg is moving up a straight road, which is inclined at an angle $\theta$ to the
horizontal, where $\sin\theta=\frac{1}{14}$ The resistance to the motion of the car from non-gravitational
forces is constant and is modelled as a single constant force of magnitude 650 N. The car’s
engine is working at a rate of 30 kW.\\
Find the acceleration of the car at the instant when its speed is 15$\text{ms}^{-1}$. }
\end{document}