### 2018-01-28 Physics Notes/Nuclear Physics/Radioactivity/Radioactivity.tex

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{tabularx}
\usetikzlibrary{decorations, decorations.text,positioning,quotes,arrows.meta,decorations.markings,3d,shapes,decorations.pathmorphing}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage[super]{nth}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\hyphenpenalty =10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Physics - Nuclear Physics}
\rfoot{Page \thepage}
\usepackage{mathtools}
\usetikzlibrary{calc,quotes,angles}
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{document}
\begin{center}
\underline{\huge Radioactivity}
\end{center}
\section{Rutherford Scattering}
\subsection{The plum pudding model}
\includegraphics[width=3cm] {plum.jpg}  
\includegraphics[width=3cm] {Plumpudding.png}
\\
The plum pudding model was the initial model of the atom, stating a sphere of positive charge with electrons embedded into it.
\subsection{Rutherford's experiment}
\includegraphics[width=6cm] {scatter.png}\\
Rutherford's experiment involved firing a beam of alpha particles at gold foil and measuring the paths of particles from the foil.
\begin{itemize}
\item Gold was used as it was expected to have a large nucleus
\item The screen fluoresces when collided with
\item This showed the atom was mostly empty space with a positive nucleus
\end{itemize}
\subsubsection{Results}

\begin{figure}[h]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Observation&Explanation\\
\hline
Most electrons pass all the way through&Atoms are mostly empty space\\
\hline
Some are deflected&The atom has a positive centre\\
\hline
Some are deflected by significant angles&The positive charge is condensed in a small area\\
\hline
\end{tabularx}
    \end{minipage}
\end{figure}
\subsection{Estimating the size of the nucleus}  
\subsubsection{Closest approach method}
\begin{tikzpicture}
\filldraw[color=red, fill=red, very thick](0,0) circle (0.2);
\draw[thick, ->,black] (0.4,0) -- (4,0) node[above,yshift=0.3cm,xshift=-2cm] {\footnotesize{Velocity reduces on approach}};
\draw[color=red, very thick](6,0) circle (1) node[] {Au};
\draw[color=black, very thick,dashed](6,0) circle (2) node[above,yshift=2cm] {EPE};
\draw[thick, ->,black] (4,0) -- (1,-1);
\end{tikzpicture}\\
\\
KE=EPE
$$8.0\times10^{-13}=\frac{1}{4\pi\epsilon_0}\times\frac{Q_{Au}}{r}\times Q_{\alpha}$$
$$r=4.55\times10^{-14}$$
\subsubsection{Estimate from scattering data}
\begin{itemize}
\item About $\frac{1}{10,000}$ deflected through more than 90\degree
\item Foil had n layers of atoms
\end{itemize}
\begin{tikzpicture}
\draw[thick, stealth-stealth,black] (6.8,-0.4) -- (6.8,0.4) node[right,midway] {d} ;
\draw[color=red, very thick](6,0) circle (0.5) node[] {\tiny{Nucleus}};
\draw[color=black, very thick](6,0) circle (2) node[above,yshift=2cm] {Atom};
\draw[thick, stealth-stealth,black] (8.5,-2) -- (8.5,2) node[right,midway] {D};
\end{tikzpicture}\\
$$\frac{\frac{1}{4}\pi d^2}{\frac{1}{4}\pi D^2}=\frac{d^2}{D^2}=\frac{1}{10,000n}$$
$n=10^4$ layers
$$\frac{d^2}{D^2}=\frac{1}{10,000\times1\times10^4}$$
$$d=\frac{D}{10,000}$$
\end{document}