### 2018-02-09 Maths Notes/Statistics/S2/Exam Questions/Continuous Distributions/Binomial/Binomial.tex
S2 Exam Qs Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{amssymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Continuous Distribution - Binomial - Exam Questions}
\end{center}
\section{Conditions}
Conditions for a binomial distribution:
\begin{itemize}
\item There are a fixed number of trials, \textbf{n}
\item There are two outcomes (success and failure)
\item Each trial is independent
\item The probability of success is constant, \textbf{p}
\end{itemize}

\section{Finding probabilities}
\textit{Bhim and Joe play each other at badminton and for each game, independently of all others, the
probability that Bhim loses is 0.2\\
Find the probability that, in 9 games:}
\subsection{P(X=?)}
\textit{Exactly 3 of the games}
\subsubsection{Method 1 - Tables}
Let X be the number of games lost
$$X\sim B(9,0.2)$$
Write the question in terms of inequalities (so tables can be used)
$$P(X=3)=P(X\leqslant3)-P(X\leqslant2)$$
Look up the values on the tables
$$0.9144-0.7382=0.1762$$
\subsubsection{Method 2 - Formula}
Let X be the number of games lost
$$X\sim B(9,0.2)$$
Write down the probability to be found:
$$P(X=3)$$
Use the formula on the data sheet:
\textcolor{red}{$$\binom{n}{x}p^x(1-p)^{n-x}$$}
Substitute values:
$$\binom{9}{3}\times(0.2)^3\times(1-0.2)^{9-3}=0.1762$$
\subsection{P(X$>$?) or P(X$<$?)}
\textit{Fewer than half the games}\\
Let X be the number of games lost
$$X\sim B(9,0.2)$$
Write down the probability to be found
$$P(X\leqslant4)$$
Look up the value on the tables
$$0.9804$$
\\
\\
If the inequality sign is the other way round subtract the probability from one.\\
\\
If the value is not on the table, use an approximation
\subsection{What to do when $p>0.5$}
Find the probability of the event not happening instead

\end{document}