### 2018-02-09 Maths Notes/Statistics/S2/Exam Questions/Continuous Distributions/Poisson/Poisson.tex

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{amssymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Continuous Distribution - Poisson - Exam Questions}
\end{center}
\section{Conditions}
The conditions for a poisson distribution are:
\begin{itemize}
\item Events occur at random
\item Events occur independently of each other
\item The average rate of occurrences remains constant
\item There is zero probability of simultaneous occurrences 
\end{itemize}
\section{Finding probabilities}
\textit{A company has a large number of regular users logging onto its website. On average 4 users
every hour fail to connect to the company's website at their first attempt.\\
Find the probability that, in a randomly chosen \textbf{2 hour} period}
\subsection{P(X=?)}
\textit{All users connect at their first attempt}
\subsubsection{Method 1 - Tables}
Let X be the number of failed connections
$$X\sim P_o(8)$$
Write the question in terms of inequalities (so tables can be used)
$$P(X=0)=P(X\leqslant0)$$
Look up the value on the tables
$$0.0002$$
\subsubsection{Method 2 - Formula}
Let X be the number failed connections
$$X\sim P_o(8)$$
Write down the probability to be found:
$$P(X=0)$$
Use the formula on the data sheet:
\textcolor{red}{$$e^{-\lambda}\frac{\lambda^x}{x!}$$}
Substitute values:
$$e^{-8}\frac{8^0}{0!}=0.0003$$
\subsection{P(X$>$?) or P(X$<$?)}
\textit{At least 4 users fail to connect at their first attempt}\\
Let X be the number of failed connections
$$X\sim P_o(8)$$
Write down the probability to be found and rewrite in terms of less than
$$P(X\geqslant4)=1-P(X\leqslant3)$$
Look up the value on the tables
$$1-0.0424=0.9576$$

\end{document}