### 2018-02-09 Maths Notes/Statistics/S2/Exam Questions/Continuous Distributions/Uniform/Uniform.tex

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{amssymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Continuous Distribution - Uniform - Exam Questions}
\end{center}
\section{Conditions}
The conditions for a uniform distribution are:
\begin{itemize}
\item All values have the same probability
\end{itemize}
\section{Finding probabilities}
The probability of any result in a uniform distribution in a range from a to b is:
$$\frac{1}{b-a}$$
\subsection{Example}
\textit{A rectangle has a perimeter of 20 cm. The length, x cm, of one side of this rectangle is
uniformly distributed between 1 cm and 7 cm.\\
Find the probability that the length of the longer side of the rectangle is more than 6 cm long}\\
Write down the distribution
$$X\sim U[1,7]$$
Find the probability wanted to be found
$$P(X<4)+P(X>6)$$
Find each probability
$$P(X>6)=1\times\frac{1}{6}=\frac{1}{6}$$
$$P(X<4)=3\times\frac{1}{6}=\frac{1}{2}$$
Add the probability to find the answer
$$\frac{1}{6}+\frac{1}{2}=\frac{2}{3}$$
\section{Mean and Variance}
\textit{Jean regularly takes a break from work to go to the post office. The amount of time Jean waits
in the queue to be served at the post office has a continuous uniform distribution between 0 and
10 minutes.}
\subsection{Mean}
\textit{Find the mean time Jean spends in the queue}\\
Write down the distribution
$$X\sim U[0,10]$$
Use the formula in the formula book
\textcolor{red}{$$\mu=\frac{1}{2}(a+b)$$}
$$\mu=\frac{1}{2}(0+10)=5$$
\subsection{Variance}
\textit{Find the variance of the time Jean spends in the queue}\\
Write down the distribution
$$X\sim U[0,10]$$
Use the formula in the formula book
\textcolor{red}{$$\textrm{Variance}=\frac{1}{12}(b-a)^2$$}
$$\textrm{Variance}=\frac{1}{12}(10-0)^2=\frac{25}{3}$$

\end{document}