### 2018-02-12 Maths Notes/Statistics/S2/Exam Questions/Discrete Distributions/Discrete Distributions.tex
Notes update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{amssymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Discrete Distributions Exam Questions}
\end{center}
\textit{A bag contains a large number of coins. It contains only 1p and 2p coins in the ratio 1:3}
\section{Populations}
\subsection{Mean and variance}
\subsubsection{Mean}
\textit{Find the mean of the population of coins}\\
Use the formula on the data sheet:
\textcolor{red}{$$\mu=\Sigma xP(X=x)$$}
Create a table of probabilities and values of x
\begin{center}
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
$x$&1p&2p\\
\hline
$P(X=x)$&$\frac{1}{4}$&$\frac{3}{4}$\\
\hline
\end{tabular}}
\end{center}
Add together the columns multiplied together
$$\mu=1\times\frac{1}{4}+2\times\frac{3}{4}=\frac{7}{4}$$
\subsubsection{Variance}
\textit{Find the variance of the population of coins}\\
Use the formula on the data sheet:
\textcolor{red}{$$\mu=\Sigma x^2P(X=x)-\mu^2$$}
Create a table of probabilities and values of x
\begin{center}
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|}
\hline
$x^2$&1&4\\
\hline
$P(X=x)$&$\frac{1}{4}$&$\frac{3}{4}$\\
\hline
\end{tabular}}
\end{center}
Add together the columns multiplied together to find $E(X^2)$
$$E(X^2)=1\times\frac{1}{4}+4\times\frac{3}{4}=\frac{13}{4}$$
Subtract find $E(X^2)-E(X)^2$ to find the variance
$$\frac{13}{4}-\Bigg(\frac{7}{4}\Bigg)^2=\frac{3}{16}$$
\newpage
\section{Samples}
\textit{List all the possible samples that can be drawn}\\
When doing this remember:
\textcolor{red}{$$\textrm{Number of samples}=\textrm{Number of options}^\textrm{Sample size}$$}
For the example:\\
Number of options=2\\
Sample size=3\\
Number of samples=$2^3$=8\\
\\
\\
List of samples:\\
(1,1,1)\\
(1,1,2)$\times$3\\
(1,2,2)$\times$3\\
(2,2,2)\\
\\
\textit{Find the sampling distribution of the mean value of these samples.}\\
\\
List the possible means
$$1,\frac{4}{3},\frac{5}{3},2$$
Calculate the probability of each, given the probability of picking a coin
$$1: \ \Bigg(\frac{1}{4}\Bigg)^3=\frac{1}{64}$$
$$\frac{4}{3}: \ 3\times\frac{1}{4}\times\frac{1}{4}\times\frac{3}{4}=\frac{9}{64}$$
$$\frac{5}{3}: \ 3\times\frac{1}{4}\times\frac{3}{4}\times\frac{3}{4}=\frac{27}{64}$$
$$2: \ \Bigg(\frac{3}{4}\Bigg)^3=\frac{27}{64}$$
Write in a table\\
\\
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|c|}
\hline
$\overline{x}$&1&$\frac{4}{3}$&$\frac{5}{3}$&2\\
\hline
$P(\overline{X}=\overline{x})$&$\frac{1}{64}$&$\frac{9}{64}$&$\frac{24}{64}$&$\frac{27}{64}$\\
\hline

\end{tabular}}

\end{document}