### 2018-02-20 Maths Notes/Mechanics/M2/Statics of rigid bodies/Statics of rigid bodies.tex
M2 Notes Update - Exam Questions

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Statics of rigid bodies}
\end{center}
\section{Equilibrium of rigid bodies}
When calculating moments, only use the component of the force acting perpendicular to the rod.\\
\\
A rigid body is in equilibrium if:
\begin{itemize}
\item The vector sum of the forces is zero
\item The sum of the moments around any point is zero
\end{itemize}
\newpage
\begin{center}
\underline{\Huge Equilibrium of rigid bodies example}
\end{center}
\includegraphics[width=8cm]{eq_rigid.png}\\
\textit{The diagram above shows a uniform rod AB of mass m and length 4a. The end A of the rod is
freely hinged to a point on a vertical wall. A particle of mass m is attached to the rod at B. One
end of a light inextensible string is attached to the rod at C, where AC = 3a. The other end of the
string is attached to the wall at D, where AD = 2a and D is vertically above A. The rod rests
horizontally in equilibrium in a vertical plane perpendicular to the wall and the tension in the
string is T. }\\
\\
\textit{Show that T = $mg\sqrt{13}$}\\
\\
Take moments about A 
$$3a\times T\cos\theta=2amg+4amg$$
Calculate $\cos\theta$ from the lengths on the diagram
$$\cos\theta=\frac{2}{\sqrt{3^2+2^2}}=\frac{2}{\sqrt{13}}$$
Substitute in the value of $\cos\theta$ and divide through by $a$
$$\frac{6}{\sqrt{13}}T=6mg$$
Multiply both sides by $\frac{\sqrt{13}}{6}$
$$T=mg\sqrt{13}$$
\\
\textit{The particle of mass m at B is removed from the rod and replaced by a particle of mass M which
is attached to the rod at B. The string breaks if the tension exceeds $2mg\sqrt{13}$. Given that the
string does not break,\\
show that $M\leqslant\frac{5}{2}m$}\\
\\
Rewrite the moments equation with the new information from the question
$$3a\times T\cos\theta=2amg+4aMg$$
Write the inequality given
$$T\leqslant 2mg\sqrt{13}$$
Substitute in the value for T
$$\frac{2mg+4Mg}{6}\sqrt{13}\leqslant2mg\sqrt{13}$$
Simplify
$$mg+2Mg\leqslant 6mg$$
$$M\leqslant\frac{5}{2}m$$
\end{document}