### 2018-03-09 Maths Notes/Statistics/S3/Goodness of fit and contingency tables/Goodness of fit and contingency tables.tex
S3 Notes Up to date

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{relsize}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{enumitem}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\hyphenpenalty=10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S3}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Goodness of fit and contingency tables}
\end{center}
Method for testing goodness of fit:
\begin{enumerate}
\item Determine which distribution would conceptually be most appropriate
\item Set significance level
\item Estimate parameters (if necessary) from observed data
\item Form hypotheses $H_0$ and $H_1$
\item Calculate expected frequencies
\item Combine expected frequencies so that none are $<5$
\item Find degrees of freedom
\item Calculate critical value of $\chi^2$ from the table
\item Calculate $\sum\frac{(O_i-E_i)^2}{E_i}$
\item See if the value is significant and draw conclusion
\end{enumerate}
$X^2$ is distributed with a chi squared distribution $\chi^2_\nu$\\
Where $\nu=$ degrees of freedom\\
$$\textrm{The number of degrees of freedom}=\textrm{Number of classes (after combining)}-1$$
\newpage
\subsection{Example}
\textit{The data in the table is thought to be modelled by a binomial $B(10,0.2)$. Use the table for the binomial cumulative distribution function to find expected values, and conduct a test to see if this is a good model. Use a 5\% significance level.}\\
\\
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
$x$&0&1&2&3&4&5&6&7&8\\
\hline
Freq of $x$&12&28&28&17&7&4&2&2&0\\
\hline
\end{tabular}
\\
\\
\textcolor{red}{Define Hypotheses}
$$H_0: \textrm{A B(10,0.2) distribution is suitable for the results}$$
$$H_1: \textrm{The distribution is not suitable for the results}$$
\textcolor{red}{Calculate the sum of frequencies}
$$N=100$$
\textcolor{red}{Complete the table of probabilities and expected frequencies, expected frequency=probability$\times$N}\\
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|X|}
\hline
$x$&0&1&2&3&4&5&6&7&8\\
\hline
$p(x)$&0.1074&0.2684&0.3020&0.2013&0.0881&0.0264&0.0055&0.0008&0.0001\\
\hline
Expected freq&10.74&26.84&30.20&20.13&8.81&2.64&0.55&0.08&0.01\\
\hline
\end{tabularx}
\\
\\
\textcolor{red}{As expected frequencies need to be greater than or equal to five, combine all probabilities greater than or equal to four}
\\
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$x$&0&1&2&3&$\geqslant4$\\
\hline
$O_i$&12&28&28&17&15\\
\hline
$E_i$&10.74&26.84&30.20&20.13&12.09\\
\hline
$\dfrac{(O_i-E_i)^2}{E_i}$&0.1478&0.0501&0.1603&0.4867&0.7004\\
\hline
\end{tabularx}
\textcolor{red}{Find the value of $\nu$}
$$\nu=5-1=4$$
\textcolor{red}{Find the value of $X^2$}
$$X^2=0.1478+0.0501+0.1603+0.4867+0.7004=1.5453$$
\textcolor{red}{Compare the value of $X^2$ to the value on the tables corresponding to the 5\% significance level and $\nu=4$}
$$9.488>1.5453$$
\textcolor{red}{Write conclusion}
\begin{center}
Not in critical region so insufficient evidence to reject $H_0$, binomial is a possible model
\end{center}

\end{document}