### 2018-03-13 Maths Notes/Statistics/S3/Goodness of fit and contingency tables/Goodness of fit and contingency tables.tex
S3 Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{relsize}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage{enumitem}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\hyphenpenalty=10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - S3}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Goodness of fit and contingency tables}
\end{center}
Method for testing goodness of fit:
\begin{enumerate}
\item Determine which distribution would conceptually be most appropriate
\item Set significance level
\item Estimate parameters (if necessary) from observed data
\item Form hypotheses $H_0$ and $H_1$
\item Calculate expected frequencies
\item Combine expected frequencies so that none are $<5$
\item Find degrees of freedom
\item Calculate critical value of $\chi^2$ from the table
\item Calculate $\sum\frac{(O_i-E_i)^2}{E_i}$
\item See if the value is significant and draw conclusion
\end{enumerate}
$X^2$ is distributed with a chi squared distribution $\chi^2_\nu$\\
Where $\nu=$ degrees of freedom\\
$$\textrm{The number of degrees of freedom}=\textrm{Number of classes (after combining)}-1$$
\newpage
\subsection{Testing a Binomial distribution as a model}
\textit{The data in the table is thought to be modelled by a binomial $B(10,0.2)$. Use the table for the binomial cumulative distribution function to find expected values, and conduct a test to see if this is a good model. Use a 5\% significance level.}\\
\\
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
$x$&0&1&2&3&4&5&6&7&8\\
\hline
Freq of $x$&12&28&28&17&7&4&2&2&0\\
\hline
\end{tabular}
\\
\\
\textcolor{red}{Define Hypotheses}
$$H_0: \textrm{A B(10,0.2) distribution is suitable for the results}$$
$$H_1: \textrm{The distribution is not suitable for the results}$$
\textcolor{red}{Calculate the sum of frequencies}
$$N=100$$
\textcolor{red}{Complete the table of probabilities and expected frequencies, expected frequency=probability$\times$N}\\
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|X|}
\hline
$x$&0&1&2&3&4&5&6&7&8\\
\hline
$p(x)$&0.1074&0.2684&0.3020&0.2013&0.0881&0.0264&0.0055&0.0008&0.0001\\
\hline
Expected freq&10.74&26.84&30.20&20.13&8.81&2.64&0.55&0.08&0.01\\
\hline
\end{tabularx}
\\
\\
\textcolor{red}{As expected frequencies need to be greater than or equal to five, combine all probabilities greater than or equal to four}
\\
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$x$&0&1&2&3&$\geqslant4$\\
\hline
$O_i$&12&28&28&17&15\\
\hline
$E_i$&10.74&26.84&30.20&20.13&12.09\\
\hline
$\dfrac{(O_i-E_i)^2}{E_i}$&0.1478&0.0501&0.1603&0.4867&0.7004\\
\hline
\end{tabularx}
\textcolor{red}{Find the value of $\nu$}
$$\nu=5-1=4$$
\textcolor{red}{Find the value of $X^2$}
$$X^2=0.1478+0.0501+0.1603+0.4867+0.7004=1.5453$$
\textcolor{red}{Compare the value of $X^2$ to the value on the tables corresponding to the 5\% significance level and $\nu=4$}
$$9.488>1.5453$$
\textcolor{red}{Write conclusion}
\begin{center}
Not in critical region so insufficient evidence to reject $H_0$, binomial is a possible model
\end{center}
\newpage
\subsection{What to do when p is not given}
\textit{A study of the number of girls in families with 5 children was done on 100 such families. The results are summarised in the following table.}\\
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Num girls(r)&0&1&2&3&4&5\\
\hline
Frequency(f)&13&18&38&20&10&1\\
\hline
\end{tabular}\\
\textit{Test, at the 5\% significance level, whether or not a binomial distribution is a good model.}\\
\\
\textcolor{red}{State hypotheses}
\begin{center}
$H_0$: The binomial distribution is a good model\\
$H_1$: The binomial distribution is not a suitable model
\end{center}
\textcolor{red}{Calculate the mean}
$$\overline{x}=\frac{0\times13+1\times18+2\times38+3\times20+4\times10+5\times1}{100}=1.99$$
\textcolor{red}{Divide the mean by n, the number of children in the families, 5, to find p.}
$$p=\frac{1.99}{5}=0.398$$
\textcolor{red}{Using the values of n and p, find the probability the value is a certain number, multiply by 100 to find the expected value.}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
\hline
r&0&1&2&3&4&5\\
\hline
p(r)&0.079&0.261&0.3456&0.229&0.0755&0.0009\\
\hline
$E_i$&7.91&26.14&34.56&22.85&7.55&0.09\\
\hline
\end{tabularx}\\
\\
\textcolor{red}{Use the values in the two above tables to find the values of $O_i$, $E_i$ and $X^2$, combine expected values when under 5}\\
\\
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
\hline
r&0&1&2&3&$>3$&Total\\
\hline
$O_i$&13&18&38&20&11&\\
\hline
$E_i$&7.91&26.14&34.56&22.85&8.54&\\
\hline
$\frac{(O_i-E_i)^2}{E_i}$&3.28&2.53&0.34&0.36&0.71&7.22\\
\hline
\end{tabularx}
\textcolor{red}{Calculate the degrees of freedom, subtracting one for a constant frequency sum and one for the estimated p}
$$\nu=5-1-1=3$$
\textcolor{red}{Find the value of $\chi^2_3$ at a 5\% significance level}
$$\chi^2_3=7.815$$
\textcolor{red}{Compare the value of $\chi^2_3$ to $X^2$ to determine the correct hypothesis}
$$7.22<7.81$$
\begin{center}
Not in critical region, so not significant, do not reject $H_0$, binomial is a suitable model
\end{center}
\newpage
\subsection{Testing a poisson distribution as a model}
\textit{The numbers of telephone calls arriving at an exchange in six-minute periods were recorded over a period of 8 hours, with the following results}\\
\\
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
Num calls(r)&0&1&2&3&4&5&6&7&8\\
\hline
Freq(f)&8&19&26&13&7&5&1&1&0\\
\hline
\end{tabular}
\\
\textit{Can these results be modelled by a Poisson distribution? Test at the 5\% significance level}\\
\\
\textcolor{red}{Calculate $\lambda$ (the mean)}
$$\lambda=\overline{x}=\frac{0\times8+1\times19+2\times26+3\times13+4\times7+5\times5+6\times1+7\times1+9\times0}{8+19+26+13+7+5+1+1+0}=2.2$$
\textcolor{red}{Use this value of $\lambda$ to find $P(r)$ and $E(r)$ by multiplying by 80}\\
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
r&$P(r)$&Expected freq of r\\
\hline
0&0.1108&8.864\\
\hline
1&0.2438&19.504\\
\hline
2&0.2681&21.448\\
\hline
3&0.1966&15.728\\
\hline
4&0.1082&8.656\\
\hline
5&0.0476&3.808\\
\hline
6&0.0174&1.392\\
\hline
$\geqslant7$&0.0075&0.6\\
\hline
\end{tabularx}
\\
\textcolor{red}{Use the two above tables to calculate $O_i$, $E_i$ and $X^2$, combining classes where needed}\\
\begin{tabularx}{\textwidth}{|X|X|X|X|}
\hline
r&$O_i$&$E_i$&$\frac{(O_i-E_i)^2}{E_i}$\\
\hline
0&8&8.864&0.0842\\
\hline
1&19&19.504&0.0130\\
\hline
2&26&21.448&0.9661\\
\hline
3&13&15.728&0.4732\\
\hline
4&7&8.656&0.3168\\
\hline
$\geqslant5$&7&5.8&0.2483\\
\hline
\end{tabularx}\\
\textcolor{red}{Find the value of $X^2$ by finding the sum of $\frac{(O_i-E_i)^2}{E_i}$}
$$X^2=0.0842+0.0130+0.9661+0.4732+0.3168+0.2483$$
\textcolor{red}{Calculate the degrees of freedom, subtracting for constant probability and estimated $\lambda$}
$$\nu=6-1-1=4$$
\textcolor{red}{Use the value for the degrees of freedom and significance level to find $\chi^2$}
$$\chi^2_4=9.488$$
\textcolor{red}{Compare the value of $\chi_4^2$ to the value of $X^2$ to determine the correct hypothesis}
$$2.1016<9.488$$
\begin{center}
Value not in critical region, non significant, accept $H_0$ insufficient evidence to reject $H_0$
\end{center}
\end{document}