### 2018-03-13 Maths Notes/Further Pure/FP2/Polar Coordinates/Polar-Coordinates.tex
FP2 Notes Update - Includes saving tikz images

\documentclass{article}
% main document, called main.tex
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{external}
\tikzexternalize[prefix=tikz/] % activate!

\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}

\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{relsize}
\pgfplotsset{width=10cm,compat=1.9}
\usepackage[super]{nth}
\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - FP2}
\rfoot{Page \thepage}

\usepackage{cancel}

\setcounter{secnumdepth}{4}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}
\titlespacing\subsubsection{0pt}{12pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}

\begin{document}
\begin{center}
\underline{\huge Polar Coordinates}
\end{center}
\begin{tikzpicture}[scale=0.7]
\begin{axis}[axis lines=middle,ymin=0,ymax=2.5,xmin=0,xmax=2.5,ticks=none]
\pgfplotsset{every tick label/.append style={font=\Huge}}

\addplot coordinates { (0,0) (2,2) };
\addplot [dashed] coordinates { (2,0) (2,2) };
\addplot [dashed] coordinates { (0,2) (2,2) };

\draw (axis cs: 0.7,0) arc[radius=70, start angle= 0, end angle= 45];
    \node[] at (axis cs: 0.5,0.2) {\huge $\theta$};
\node[blue] at (axis cs: 1,1.2) {\huge $r$};
\node[black] at (axis cs: 2.1,1) {\huge $y$};
\node[black] at (axis cs: 1,2.1) {\huge $x$};
\end{axis}
\end{tikzpicture}\\
Usual conventions are either $-\pi<\theta\leqslant\pi$ or $0\leqslant\theta<\pi$\\
\\
$x=r\cos\theta$\\
$y=r\sin\theta$\\
$r^2=x^2+y^2$\\
$\theta=\arctan\Big(\dfrac{y}{x}\Big)$
\section{Converting between polar and Cartesian form}
\subsection{Example 1}
\textit{Find the Cartesian equation of:}
$r=5$
$$\sqrt{x^2+y^2}=5$$
$$x^2+y^2=25$$
\subsection{Example 2}
\textit{Find the Cartesian equation of:}
$$r=2+\cos2\theta$$
Replace r and convert $\cos2\theta$
$$\sqrt{x^2+y^2}=2+\cos^2\theta-\sin^2\theta$$
Convert $\sin^2\theta$ and $\cos^2\theta$
$$\sqrt{x^2+y^2}=2+\frac{x^2}{x^2+y^2}-\frac{y^2}{x^2+y^2}$$
Multiply all terms by $x^2+y^2$
$$(x^2+y^2)^{\frac{3}{2}}=3x^2+y^2$$
\newpage
\section{Sketching polar curves}
To plot less standard types we look for axes intercepts and max and min values
\subsection{Standard types}
\subsubsection{r=a}
A circle, centre (0,0) with a radius of a
\subsubsection{$\theta=\alpha$}
A half line starting from (0,0) making an angle of $\alpha$ with the initial line (positive x axis)
\subsubsection{$r=a\theta$}
A spiral starting at the origin
\subsection{Cardioid type}
\subsubsection{$r=a(1+\cos\theta)$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=360,smooth] (x,{6*(1+cos(x))});
\end{axis}
\end{tikzpicture}
\subsubsection{$r=a(p+q\cos\theta)$}
\paragraph{p=q}
Factor out the value of p and plot as normal
\paragraph{$p\geqslant2q$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=360,smooth] (x,{(5+2*cos(x))});
\end{axis}
\end{tikzpicture}
\newpage
\paragraph{$q\leqslant p<2q$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=360,smooth] (x,{(3+2*cos(x))});
\end{axis}
\end{tikzpicture}
\subsubsection{$r^2=a^2\cos2\theta$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=8000] (x,{sqrt(cos(2*x))});
\end{axis}
\end{tikzpicture}
\\
The 4 asymptotes are half lines at $\theta=\frac{\pi}{4},\frac{3\pi}{4},\frac{5\pi}{4},\frac{7\pi}{4}$
\subsection{$r^2=a^2\sin2\theta$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=8000] (x,{sqrt(sin(2*x))});
\end{axis}
\end{tikzpicture}
\\
This is an anticlockwise rotation of $r^2=a^2\cos2\theta$ by $\frac{\pi}{4}$, it is this not $\frac{\pi}{2}$ as the phase difference between $\cos2\theta$ and $\sin2\theta$ is $\frac{\pi}{4}$ compared to $\frac{\pi}{2}$ with one $\theta$.
\subsubsection{$r=a(p+q\sin\theta)$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
    ticks=none
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=360,smooth] (x,{6*(1+sin(x))});
\end{axis}
\end{tikzpicture}
\\
This is a rotation of $r=a(p+q\cos\theta)$ anticlockwise by $\frac{\pi}{2}$
\subsection{Other types}
\subsubsection{$r=a\sin3\theta$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
    axis equal image,
    enlargelimits=true,
     ]
    \addplot[data cs=polar,red,domain=0:360,samples=360,smooth] (x,{sin(3*x)});
\end{axis}
\end{tikzpicture}
\\
Asymptotes occur when $r=0$
\subsubsection{$r=2\sec(\theta-\frac{\pi}{3})$}
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    axis lines=center,
     ]
    \addplot[data cs=polar,red,domain=-10:100,samples=360,smooth] (x,{2/(cos(x-60))});
\end{axis}
\end{tikzpicture}

\end{document}