### 2018-04-06 Maths Notes/Statistics/S2/Exam Questions/Continuous Distributions/Normal/Normal.tex

\documentclass{article}[18pt]
\usepackage{/home/sam/Documents/School_Notes/format}
\lhead{A Level Maths - S2}

\begin{document}
\begin{center}
\underline{\huge Continuous Distribution - Normal- Exam Questions}
\end{center}
\section{Approximations}
\subsection{Poisson}
\subsubsection{Cases where normal can be used to approximate poisson}
\begin{itemize}
\item $\lambda>10$
\end{itemize}
\subsubsection{Conversion}
$$X\sim P_o(\lambda)$$
$$X\sim N(\lambda,\lambda)$$
\subsubsection{Continuity correction}
As the poisson distribution is discrete and the normal distribution is continuous continuity correction must be used.\\
\\
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|}
\hline
Probability&Corrected Probability\\
\hline
$P(X=n)$&$P(n-0.5<X<n+0.5)$\\
\hline
$P(X>n)$&$P(X>n+0.5)$\\
\hline
$P(X\leqslant n)$ &$P(X<n+0.5)$\\
\hline
$P(X<n)$&$P(X<n-0.5)$\\
\hline
$P(X\geqslant n)$&$P(X>n-0.5)$\\
\hline
\end{tabular}}
\subsubsection{Example question}
\textit{During the summer the mean number of yachts hired per week is 25.}\\
\textit{The company has only 30 yachts for hire.}\\
\textit{Using a suitable approximation find the probability that the demand for yachts cannot be
met on a particular Saturday in the summer.}\\
\\
Let X Represent the number of yachts hired
$$X\sim P_o(25)$$
$$X\sim N(25,25)$$
Write inequality
$$P(X>30)$$
Apply continuity correction
$$P(X>30.5)$$
Do Z transformation
$$Z=\frac{30.5-25}{5}=1.1$$
Lookup value in tables
$$0.8643$$
Subtract from one as probability is greater than
$$1-0.8643=0.1357$$
\newpage
\subsection{Binomial}
\subsubsection{Conditions for normal to approximate binomial}
\begin{itemize}
\item $n>10$
\item $p<0.5$
\end{itemize}
\subsubsection{Conversion}
$$X\sim B(n,p)$$
$$X\sim N(np,np(1-p))$$
\subsubsection{Continuity correction}
Binomial will also require continuity correction when approximated to normal as it is discrete and normal is continuous.\\
\\
{\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|}
\hline
Probability&Corrected Probability\\
\hline
$P(X=n)$&$P(n-0.5<X<n+0.5)$\\
\hline
$P(X>n)$&$P(X>n+0.5)$\\
\hline
$P(X\leqslant n)$ &$P(X<n+0.5)$\\
\hline
$P(X<n)$&$P(X<n-0.5)$\\
\hline
$P(X\geqslant n)$&$P(X>n-0.5)$\\
\hline
\end{tabular}}
\subsubsection{Example Question}
\textit{In a large college 58\% of students are female and 42\% are male.\\ A random sample of 100
students is chosen from the college.\\ Using a suitable approximation find the probability that
more than half the sample are female.}
Write the question in the binomial for females
$$X\sim B(100,0.58)$$
Use the conversion formula to convert to normal
$$X\sim N(58,24.36)$$
Write the inequality then apply continuity correction
$$P(X>50)\rightarrow P(X>50.5)$$
Find the Z value
$$Z=\frac{50.5-58}{\sqrt{24.36}}\approx-1.52$$
Convert the Z value to a probability using the tables
$$p=0.9357$$
\newpage

\end{document}