### 2018-04-06 Maths Notes/Statistics/S2/Exam Questions/Continuous Distributions/Uniform/Uniform.tex

\documentclass{article}[18pt]
\usepackage{/home/sam/Documents/School_Notes/format}
\lhead{A Level Maths - S2}

\begin{document}
\begin{center}
\underline{\huge Continuous Distribution - Uniform - Exam Questions}
\end{center}
\section{Conditions}
The conditions for a uniform distribution are:
\begin{itemize}
\item All values have the same probability
\end{itemize}
\section{Finding probabilities}
The probability of any result in a uniform distribution in a range from a to b is:
$$\frac{1}{b-a}$$
\subsection{Example}
\textit{A rectangle has a perimeter of 20 cm. The length, x cm, of one side of this rectangle is
uniformly distributed between 1 cm and 7 cm.\\
Find the probability that the length of the longer side of the rectangle is more than 6 cm long}\\
Write down the distribution
$$X\sim U[1,7]$$
Find the probability wanted to be found
$$P(X<4)+P(X>6)$$
Find each probability
$$P(X>6)=1\times\frac{1}{6}=\frac{1}{6}$$
$$P(X<4)=3\times\frac{1}{6}=\frac{1}{2}$$
Add the probability to find the answer
$$\frac{1}{6}+\frac{1}{2}=\frac{2}{3}$$
\section{Mean and Variance}
\textit{Jean regularly takes a break from work to go to the post office. The amount of time Jean waits
in the queue to be served at the post office has a continuous uniform distribution between 0 and
10 minutes.}
\subsection{Mean}
\textit{Find the mean time Jean spends in the queue}\\
Write down the distribution
$$X\sim U[0,10]$$
Use the formula in the formula book
\textcolor{red}{$$\mu=\frac{1}{2}(a+b)$$}
$$\mu=\frac{1}{2}(0+10)=5$$
\subsection{Variance}
\textit{Find the variance of the time Jean spends in the queue}\\
Write down the distribution
$$X\sim U[0,10]$$
Use the formula in the formula book
\textcolor{red}{$$\textrm{Variance}=\frac{1}{12}(b-a)^2$$}
$$\textrm{Variance}=\frac{1}{12}(10-0)^2=\frac{25}{3}$$
\newpage
\subsubsection{Proof of variance formula}
$$E(X)=\frac{1}{2}(a+b)$$
$$E(X^2)=\frac{1}{b-a}\int^b_a x^2\ dx$$
$$E(X^2)=\frac{1}{b-a}\Bigg[\frac{x^3}{3}\Bigg]^b_a$$
$$E(X^2)=\frac{b^3-a^3}{b-a}$$
$$E(X^2)=\frac{(b-a)(b^2+a^2+ab}{3(b-a)}$$
$$E(X^2)=\frac{b^2+a^2+ab}{3}$$
$$Var(X)=\frac{b^2+a^2+ab}{3}-\Big(\frac{a+b}{2}\Big)^2$$
$$Var(X)=\frac{b^2+a^2+ab}{3}-\frac{a^2+2ab+b^2}{4}$$
$$Var(X)=\frac{b^2-2ab+a^2}{12}$$
$$Var(X)=\frac{(b-a)^2}{12}$$

\end{document}