### 2018-04-06 Maths Notes/Statistics/S2/Exam Questions/Hypothesis tests/Binomial/Binomial.tex

\documentclass{article}[18pt]
\usepackage{/home/sam/Documents/School_Notes/format}
\lhead{A Level Maths - S2}

\begin{document}
\begin{center}
\underline{\huge Hypothesis tests - Binomial - Exam Questions}
\end{center}
\textit{A company claims that a quarter of the bolts sent to them are faulty. To test this claim the
number of faulty bolts in a random sample of 50 is recorded.}
\section{Finding the critical region}
\textit{Using a 5\% significance level, find the critical region for a two-tailed test of the
hypothesis that the probability of a bolt being faulty is $\frac{1}{4}$. The probability of rejection in
either tail should be as close as possible to 0.025}\\
\\
Write down the distribution:
$X\sim B(50,0.25)$
For the lower value, look for the value closest to the probability of rejection
$$P(X\leqslant6)=0.0194$$
To find the upper value look for the value closest to (1-probability of rejection). Then add one to this value as:
\textcolor{red}{$$P(X\geqslant c)=1-P(X\leqslant c-1)$$}
$$P(X\leqslant18)=0.9713$$
$$P(X\geqslant19)=0.0287$$
\begin{center}
The critical values are 6 and 19
\end{center}
\textit{Find the actual significance}\\
The actual significance is the true probability of rejection
$$0.0194+0.0287=0.0481$$
\section{One tailed test}
\textit{The machine making the bolts was reset and another sample of 50 bolts was taken. Only 5 were
found to be faulty}\\
\\
\textit{Test at the 1\% level of significance whether or not the probability of a faulty bolt has
decreased. State your hypotheses clearly.}\\
\\
State the null and alternative hypothesis related to the $\frac{1}{4}$ probability mentioned earlier in the question
$$H_0: \ p=0.25$$
$$H_1: \ p<0.25$$
Write a probability based on the information given in the question
$$P(X\leqslant5)$$
Look up the probability on the tables
$$P(X\leqslant5)=0.007$$
Compare this to the significance level
$$0.007<0.01$$
Write the conclusion
\begin{center}
5 is in the critical region, reject $H_0$, the test is significant, there is evidence that the probability of faulty bolts has decreased.
\end{center}
\newpage
\section{Two tailed test}
\textit{A teacher thinks that 20\% of the pupils in a school read the Deano comic regularly.\\
\\
He chooses 20 pupils at random and finds 9 of them read Deano.}\\
\\
\textit{Test, at the 5\% level of significance, whether or not there is evidence that the
percentage of pupils that read Deano is different from 20\%. State your hypotheses
clearly.}\\
\\
Write down the null and alternate hypothesis
$$H_0: \ p=0.2$$
$$H_1: \ p\neq0.2$$
Write down the distribution
$$X\sim B(20,0.2)$$

Write down a probability based on information in the question
$$P(X\geqslant9)$$
Find this probability using the tables
$$P(X\geqslant9)=1-P(X\leqslant8)=1-0.99=0.01$$
Compare this value with the significance level, remembering to split it
$$0.01<0.025$$
Write conclusion
\begin{center}
Probability is smaller than significance level, reject $H_0$ accept $H_1$, there is evidence that the percentage of pupils that read beano is not 20\%. 
\end{center}
\end{document}