### 2018-04-06 Maths Notes/Statistics/S3/Exam Questions/Confidence Intervals/Confidence Intervals.tex

\documentclass{article}[18pt]
\usepackage{/home/sam/Documents/School_Notes/format}
\lhead{A Level Maths - S3}

\begin{document}
\begin{center}
\underline{\huge Exam Questions - Confidence Intervals}
\end{center}
\section{Normal mean}
\textit{A woodwork teacher measures the width, w mm, of a board. The measured width, x mm, is
normally distributed with mean w mm and standard deviation 0.5 mm.\\
The same board is measured 16 times and the results are recorded.\\
Given that the mean of these 16 measurements is 35.6 mm,\\
find a 98\% confidence interval for w.}
\textcolor{red}{Find the standard error}
$$\textrm{Standard error}=\frac{\sigma}{\sqrt{n}}=\frac{0.5}{\sqrt{16}}=\frac{1}{8}$$
\textcolor{red}{Find the z greater value for 1\%}
$$P(Z>2.3263)=0.01$$
\textcolor{red}{Substitute in the values}
$$\overline{x}\pm z\times\textrm{Standard error}$$
$$35.6\pm2.3263\times\frac{1}{8}$$

\end{document}