### 2018-04-06 Maths Notes/Statistics/S2/Definitions/Definitions.tex
S2/S3 Definitions update

\documentclass{article}[18pt]
\usepackage{/home/sam/Documents/School_Notes/format}
\lhead{A Level Maths - S2}

\begin{document}
\begin{center}
\underline{\huge Definitions}
\end{center}
\large
\section{Population and Samples}
Statistic - A random variable, a function of known observations from a population, no unknown parameters\\
\\
Sampling unit - An individual unit of the population\\
\\
Sampling frame - A list of all sampling units\\
\\
Sampling distribution - All possible samples are chosen from the population\\
\\
Census - When every member of the population is investigated\\
\\
Population - A collection of all items\\
\\
Finite Population - One in which each individual member can be given a number\\
\\
Infinite population - One where it is impossible to number each individual member\\
\\
Sample - A selection of items from a population\\
\\
Bias - Anything which prevents a sample from being truly representative\\
\\
\section{Hypothesis Tests}
Significance level - The probability of incorrectly rejecting $H_0$\\
\\
Hypothesis test - A mathematical procedure to examine a value of
a population parameter proposed by the null hypothesis compared with an alternative hypothesis\\
\\
Critical Region - The range of values that would lead to the rejection of $H_0$, and so the acceptance of $H_1$\\
\\
Hypothesis - A statement concerning a population parameter\\
\\
Test Statistic - Data collected to test a hypothesis
\end{document}