### 2018-04-10 Maths Notes/Further Pure/FP2/Exam Questions/Further Complex Numbers/Complex Numbers Exam.tex
FP2 Notes Update

\documentclass{article}[18pt]
\usepackage{../../../../../format}
\lhead{A Level Maths - FP2}

\begin{document}
\begin{center}
\underline{\huge Complex Numbers - Exam Questions}
\end{center}
\section{Example 1 - Finding roots of $z^a$}
\textit{Find the four roots of the equation {\large$z^4=8(\sqrt{3}+i)$} in the form {\large$z=re^{i\theta}$}}
\\
\textcolor{red}{Convert into the form {\large $z^4=re^{i\theta}$}}
$$\theta=\arctan\bigg(\frac{1}{\sqrt{3}}\bigg)=\frac{\pi}{6}$$
$$r=\sqrt{(8\sqrt{3})^2+8^2}=16$$
{\Large
$$z^4=16e^{i(\frac{\pi}{6}+2k\pi)}$$}
\textcolor{red}{Take the fourth root of both sides}
{\Large
$$z=2e^{i(\frac{\pi}{24}+\frac{1}{2}k\pi)}$$}
\textcolor{red}{Substitute in values of k to get values of $\theta$ between 0 and $2\pi$}
{\large
$$z=2e^{i\frac{\pi}{24}}$$
$$z=2e^{i\frac{13\pi}{24}}$$
$$z=2e^{i\frac{25\pi}{24}}$$
$$z=2e^{i\frac{37\pi}{24}}$$}

\end{document}