### 2018-04-13 Maths Notes/Statistics/S3/Any kind of hypothesis test in S3/S3 Hypothesis Tests.tex
Notes Update

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - S3}

\begin{document}
\begin{center}
\underline{\huge What the hypotheses are for all S3 tests}
\end{center}
\section{Estimation, confidence intervals and tests}
\subsection{The difference between means}
$$H_0:\mu_a=\mu_b$$
$$H_1:\mu_a>\mu_b \ \textrm{or} \ 
\mu_a<\mu_b \  \textrm{or} \ 
\mu_a\neq\mu_b$$
\section{Goodness of fit and contingency tables}
\subsection{Goodness of fit of distributions}
\begin{center}
$H_0$: The distribution is suitable for the results\\
$H_1$: The distribution is not suitable for the results
\end{center}
\subsection{Contingency tables}
\begin{center}
$H_0$: The two data sets are independent\\
$H_1$: The two data sets are not independent
\end{center}
\section{Regression and correlation}
\subsection{Spearman's rank}
\begin{center}
$H_0$: $\rho_s=0$\\
$H_1$: $\rho_s<0$ or $\rho_s>0$ or $\rho_s\neq0$
\end{center}

\end{document}