### 2018-04-19 Maths Notes/Statistics/S3/Overall Notes/S3 Notes.tex

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - S3}

\begin{document}
\begin{center}
\underline{\huge S3 Notes}
\end{center}
\section{Estimation, confidence intervals and tests}
\subsection{Estimators}
Bias is:
$$E(T)-\theta$$
To make the formula for $S^2$ in the formula book easier to use replace the top of the fraction with $S_{xx}$ from the formula book
\subsection{Standard error}
Standard error is:
$$\frac{\sigma}{\sqrt{n}} \ \textrm{or} \ \frac{s}{\sqrt{n}}$$
This can be remembered as the square root of the approximated variance given in the formula book
\subsection{Central limit theorem}
CLT states that sample means from a distribution where n is greater than 50 is:
$$\overline{X}\sim N(\mu,\frac{\sigma^2}{n})$$
\subsection{Confidence intervals}
To calculate the confidence interval use:
$$\overline{x}\pm z\times\frac{\sigma}{\sqrt{n}}$$
Where z is the value from the percentage points table that correlates to the probability in each of the tails on the normal distribution
\section{Goodness of fit and Contingency tables}
\subsection{Goodness of fit method}
Method for testing goodness of fit:
\begin{enumerate}
\item Determine which distribution would conceptually be most appropriate
\item Set significance level
\item Estimate parameters (if necessary) from observed data
\item Form hypotheses $H_0$ and $H_1$
\item Calculate expected frequencies
\item Combine expected frequencies so that none are $<5$
\item Find degrees of freedom
\item Calculate critical value of $\chi^2$ from the table
\item Calculate $\sum\frac{(O_i-E_i)^2}{E_i}$
\item See if the value is significant and draw conclusion
\end{enumerate}
$X^2$ is distributed with a chi squared distribution $\chi^2_\nu$\\
Where $\nu=$ degrees of freedom\\
$$\textrm{The number of degrees of freedom}=\textrm{Number of classes (after combining)}-1$$
\subsection{Contingency tables}
\begin{itemize}
\item We use this test to see if two factors are independent of each other
\item We describe them by: Number of rows $\times$ Number of columns
\item $H_0$ is that they are independent
\item $H_1$ is that they are not independent
\item $$\textrm{Expected values}=\frac{\textrm{Row total}\times\textrm{Column total}}{\textrm{Grand total}}$$
\item $$\nu=(\textrm{Number of rows-1})(\textrm{Number of columns-1})$$
\end{itemize}
\section{Combinations of normal distributions}
\subsection{Sums and Differences}
$$\mu_{x+y}=\mu_x+\mu_y$$
$$\mu_{x-y}=\mu_x-\mu_y$$
$$\sigma^2_{x+y}=\sigma^2_x+\sigma^2_y$$
$$\sigma^2_{x-y}=\sigma^2_x+\sigma^2_y$$
\subsection{Multiples}
$$\mu_{ax+by}=a\mu_x+b\mu_y$$
$$\mu_{ax-by}=a\mu_x-b\mu_y$$
$$\sigma^2_{ax+by}=a^2\sigma^2_x+b^2\sigma^2_x$$
$$\sigma^2_{ax-by}=a^2\sigma^2_x+b^2\sigma^2_x$$
\subsection{Addition vs Multiplication}
$$E(X_1+X_2)=E(2X_1)$$
However
$$Var(X_1+X_2)\neq Var(2X_1)$$
\\
$$Var(X_1+X_2)=\sigma_1^2+\sigma_1^2=2\sigma_2^2$$
$$Var(2X_1)=4\sigma^2_1$$
\section{Regression and Correlation}
Spearman's rank: The tendency for y to increase as x increases\\
PMCC:The closeness of the data to follow linear relationship\\
\\
For example an $x^3$ graph would have a Spearman's rank of 1 as it always increases, however the PMCC would not be 1 as it does not follow a linear relationship.

\end{document}