### 2018-05-21 Test/Definitions/Definitions.tex

\documentclass{article}
\usepackage[a4paper,hmargin=3cm,vmargin=5cm]{geometry}
\usepackage{pdflscape,lipsum,graphicx}
\usepackage[absolute]{textpos}
\usepackage{fancyhdr}
\usepackage{afterpage}

\begin{document}
Test
\newpage

\newgeometry{landscape}     % use whatever margins you want for left, right, top and bottom.
...
... %<The contents of enlarged page(s)>
...    
\restoregeometry     %so it does not affect the rest of the pages.
\end{document}