### 2018-05-27 Maths Notes/Core/C4/Cheat Sheet/C4 Cheat Sheet.tex
Notes update

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - C4}

%File specific Preamble
\usepackage{polynom} %Provides long division

\begin{document}
\begin{center}
\underline{\huge C4 Cheat Sheet}
\end{center}
\section{Partial fractions}
\subsection{Split a fraction whose denominator is a product of linear expressions}
Set the fraction equal to the sum of constants over each linear expression, then multiply by the denominator and solve by substitution.
\subsection{Split a fraction where one or more of the factors in the denominator are squared}
Each squared term gets two terms in the summation, one of the non squared term, and one of the squared term.
\subsection{Deal with top heavy fractions}
Use long division to simplify, for example simplifying $\dfrac{3x^2-3x-2}{(x-1)(x-2)}$\\
\textbf{Long division to find remainder}\\
$\polylongdiv{3x^2-3x-2}{x^2-3x+2}$\\
\\
\textbf{Re-write with remainder}\\
\\
$3+\dfrac{6x-8}{(x-1)(x-2)}$\\
\textbf{Partial fractions can then be applied normally}
\section{Parametric equations}

\end{document}